### 2021-10-01 RFCs/HDF5/VOL/connector_author_guide/connector_interface.tex

\begin{lstlisting}[caption={VOL connector class, H5VLpublic.h},captionpos=b, label={lst:VOLclass}]
/* Class information for each VOL driver */
typedef struct H5VL_class_t {
    /* Overall connector fields & callbacks */                                   
    unsigned int version;                   /* VOL connector class struct version #     */
    H5VL_class_value_t value;               /* Value to identify connector              */
    const char *name;                       /* Connector name (MUST be unique!)         */
    unsigned cap_flags;                     /* Capability flags for connector           */
    herr_t (*initialize)(hid_t vipl_id);    /* Connector initialization callback        */
    herr_t (*terminate)(void);              /* Connector termination callback           */

    /* VOL framework */                                                          
    H5VL_info_class_t       info_cls;       /* VOL info fields & callbacks  */   
    H5VL_wrap_class_t       wrap_cls;       /* VOL object wrap / retrieval callbacks */

    /* Data Model */                                                             
    H5VL_attr_class_t       attr_cls;       /* Attribute (H5A*) class callbacks */
    H5VL_dataset_class_t    dataset_cls;    /* Dataset (H5D*) class callbacks   */
    H5VL_datatype_class_t   datatype_cls;   /* Datatype (H5T*) class callbacks  */
    H5VL_file_class_t       file_cls;       /* File (H5F*) class callbacks      */
    H5VL_group_class_t      group_cls;      /* Group (H5G*) class callbacks     */
    H5VL_link_class_t       link_cls;       /* Link (H5L*) class callbacks      */
    H5VL_object_class_t     object_cls;     /* Object (H5O*) class callbacks    */

    /* Infrastructure / Services */
    H5VL_introspect_class_t introspect_cls; /* Container/connector introspection class callbacks */
    H5VL_request_class_t    request_cls;    /* Asynchronous request class callbacks */
    H5VL_blob_class_t       blob_cls;       /* 'blob' callbacks */
    H5VL_token_class_t      token_cls;      /* VOL connector object token class callbacks */

    /* Catch-all */                                                              
    herr_t (*optional)(void *obj, int op_type, hid_t dxpl_id, void **req, va_list arguments); /* Optional callback */
} H5VL_class_t;
\end{lstlisting}

@@

The \texttt{value} field is a unique integer identifier that should be greater than 256 for external, non-library connectors. Setting it in the VOL structure is required.

The \texttt{name} field is a string that uniquely identifies the VOL connector name. Setting it in the VOL structure is required.

The \texttt{cap\_flags} field will be used to hold bitwise capability/feature flags that can be used with a future query API call to determine which operations and capabilities are supported by a the VOL connector. This feature is currently under development.

@@

The \texttt{terminate} field is a function pointer to a routine that a connector implements to terminate or finalize access to the connector. Implementing this function by the connector is not required since some connectors do not require any termination phase to the connector. In that case, the value of the function pointer should be set to NULL. Connector specific variables that are required to be passed from users should be passed through the VOL terminate property list. Generic properties can be added to this property class for user-defined connectors that cannot modify the HDF5 library to add internal properties. For more information consult the property list reference manual pages.

The rest of the fields in the \texttt{H5VL\_class\_t} struct are "subclasses'' that define all the VOL function callbacks that
are mapped to from the HDF5 API layer. Those subclasses are categorized into two categories, Data Model and Services. Data Model classes are those that provide functionality for accessing an HDF5 container and objects in that container as defined by the HDF5 data model.  Service classes are those that provide services for users that are not related to the data model specifically. Asynchronous operations, for example, are a service that most connectors can implement, so we add a class for it in the VOL structure. If a service becomes generic enough and common among many connectors, a class for it should be added to the VOL structure. However, many connectors can/will provide services that are not shared by other connectors. A good way to support these services is through an optional callback in the VOL structure which can be a hook from the API to the connector that provides those services, passing any necessary arguments needed without the HDF5 library having to worry about supporting that service. A similar API operation to allow users to use that service will be added. This API call would be similar to an “ioctl” call where any kind of operation can be supported and passed down to the connector that has enough knowledge from the user to interpret the type of the operation. All classes and their defined callbacks will be detailed in the following sub-sections.

To handle that large set of API routines, each class in the Data Model category has three generic callbacks, \texttt{get}, \texttt{specific}, and \texttt{optional} to handle the three set of API operations outline above respectively. The callbacks will have a \texttt{va\_list} argument to handle the different set of parameters that could be passed in.  The \texttt{get} and \texttt{specific} callbacks also have an \texttt{op\_type} that contain an \texttt{enum} of the type of operation that is requested to be performed. Using that type, the \texttt{va\_list} argument can be parsed. The \texttt{optional} callback is a free for all callback where anything from the API layer is passed in directly. This callback is used to support connector specific operations in the API that other connectors should or would not know about. More information about types and the arguments for each type will be detailed in the corresponding class arguments.

@@

\begin{lstlisting}[caption={Structure for attribute callback routines, H5VLconnector.h}, captionpos=b, label={lst:ATTRclass}]
typedef struct H5VL_attr_class_t {                                               
    void *(*create)(void *obj, const H5VL_loc_params_t *loc_params, const char *attr_name, hid_t type_id, hid_t space_id, hid_t acpl_id, hid_t aapl_id, hid_t dxpl_id, void **req);
    void *(*open)(void *obj, const H5VL_loc_params_t *loc_params, const char *attr_name, hid_t aapl_id, hid_t dxpl_id, void **req);
    herr_t (*read)(void *attr, hid_t mem_type_id, void *buf, hid_t dxpl_id, void **req);
    herr_t (*write)(void *attr, hid_t mem_type_id, const void *buf, hid_t dxpl_id, void **req);
    herr_t (*get)(void *obj, H5VL_attr_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
    herr_t (*specific)(void *obj, const H5VL_loc_params_t *loc_params, H5VL_attr_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);            
    herr_t (*optional)(void *obj, H5VL_attr_optional_t opt_type, hid_t dxpl_id, void **req, va_list arguments); 
    herr_t (*close) (void *attr, hid_t dxpl_id, void **req);                     
} H5VL_attr_class_t; 
\end{lstlisting}

@@

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
    herr_t (*get)(void *obj, H5VL_attr_get_t get_type, hid_t dxpl_id, 
        void **req, va_list arguments);
\end{lstlisting}

The \texttt{get\_type} argument is an \texttt{enum}:
\begin{lstlisting}
/* types for attribute GET callback */
typedef enum H5VL_attr_get_t {                                                   
    H5VL_ATTR_GET_ACPL,                     /* creation property list              */
    H5VL_ATTR_GET_INFO,                     /* info                                */
    H5VL_ATTR_GET_NAME,                     /* access property list                */
    H5VL_ATTR_GET_SPACE,                    /* dataspace                           */
    H5VL_ATTR_GET_STORAGE_SIZE,             /* storage size                        */
    H5VL_ATTR_GET_TYPE                      /* datatype                            */
} H5VL_attr_get_t; 
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{13.5cm}}
  \texttt{obj} & (IN): An attribute or location object where information
  needs to be retrieved from.\\
  \texttt{get\_type} & (IN): The type of the information to retrieve.\\
  \texttt{dxpl\_id} & (IN): The data transfer property list.\\
  \texttt{req} & (IN/OUT): A pointer to the asynchronous request of the
  operation created by the connector.\\
  \texttt{arguments} & (IN/OUT): argument list containing parameters and
  output pointers for the get operation. \\
\end{tabular}
\end{mdframed}

The \texttt{arguments} argument contains a variable list of arguments
depending on the \texttt{get\_type} parameter. The following list shows
the argument list, in order, for each type:

\begin{itemize}

\item \texttt{H5VL\_ATTR\_GET\_ACPL}, to retrieve the attribute creation
  property list of the attribute specified in \texttt{obj}:
  \begin{enumerate}
  \item \texttt{hid\_t *ret\_id} (OUT): buffer for the identifier of the
    attribute creation property list.
  \end{enumerate}

\item \texttt{H5VL\_ATTR\_GET\_INFO}, to retrieve the attribute info:
  \begin{enumerate}
  \item \texttt{H5VL\_loc\_params\_t *loc\_params} (IN): Pointer to the location parameters
    explained in Section~\ref{sec:map}. 
  \item \texttt{H5A\_info\_t *ainfo} (OUT): info structure to fill the
    attribute info in.
  \end{enumerate}

\item \texttt{H5VL\_ATTR\_GET\_NAME}, to retrieve an attribute name on a
  particular object specified in \texttt{obj}:
  \begin{enumerate}
  \item \texttt{H5VL\_loc\_params\_t *loc\_params} (IN): Pointer to the location parameters
    explained in Section~\ref{sec:map}. The type could be either
    \texttt{H5VL\_OBJECT\_BY\_SELF} meaning \texttt{obj} is the attribute,
    or \texttt{H5VL\_OBJECT\_BY\_IDX} meaning the attribute to retrieve
    the name for should be looked up using the index information on
    the object in \texttt{obj} and the index information in \texttt{loc\_params}.
  \item \texttt{size\_t buf\_size} (IN): the size of the buffer to store
    the name in.
  \item \texttt{void *buf} (OUT): Buffer to store the name in.
  \item \texttt{ssize\_t *ret\_val} (OUT): return the actual size needed
    to store the fill attribute name.
  \end{enumerate}

\item \texttt{H5VL\_ATTR\_GET\_SPACE}, to retrieve the dataspace of the
  attribute specified in \texttt{obj}:
  \begin{enumerate}
  \item \texttt{hid\_t *ret\_id} (OUT): buffer for the identifier of the
    attribute dataspace.
  \end{enumerate}

\item \texttt{H5VL\_ATTR\_GET\_STORAGE\_SIZE}, to retrieve the storage
  size of the attribute specified in \texttt{obj}:
  \begin{enumerate}
  \item \texttt{hsize\_t *ret} (OUT): buffer for the storage size of
    the attribute in the container.
  \end{enumerate}

\item \texttt{H5VL\_ATTR\_GET\_TYPE}, to retrieve the datatype of the
  attribute specified in \texttt{obj}:
  \begin{enumerate}
  \item \texttt{hid\_t *ret\_id} (OUT): buffer for the identifier of the
    attribute datatype.
  \end{enumerate}

\end{itemize}

@@

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
	herr_t (*specific)(void *obj, H5VL_loc_params_t *loc_params, H5VL_attr_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
\end{lstlisting}

The \texttt{specific\_type} argument is an \texttt{enum}:
\begin{lstlisting}
/* types for attribute SPECFIC callback */
typedef enum H5VL_attr_specific_t {                                              
    H5VL_ATTR_DELETE,                       /* H5Adelete(_by_name/idx)             */
    H5VL_ATTR_EXISTS,                       /* H5Aexists(_by_name)                 */
    H5VL_ATTR_ITER,                         /* H5Aiterate(_by_name)                */
    H5VL_ATTR_RENAME                        /* H5Arename(_by_name)                 */
} H5VL_attr_specific_t; 
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{13.5cm}}
  \texttt{obj} & (IN): The location object  where the operation needs to happen.\\
  \texttt{loc\_params} & (IN): A pointer to the location parameters as explained in Section~\ref{sec:map}.\\
  \texttt{specific\_type} & (IN): The type of the operation.\\
  \texttt{dxpl\_id} & (IN): The data transfer property list.\\
  \texttt{req} & (IN/OUT): A pointer to the asynchronous request of the
  operation created by the connector.\\
  \texttt{arguments} & (IN/OUT): argument list containing parameters and
  output pointers for the get operation. \\
\end{tabular}
\end{mdframed}

The \texttt{arguments} argument contains a variable list of arguments
depending on the \texttt{specific\_type} parameter. The following list shows
the argument list, in order, for each type:

\begin{itemize}
\item \texttt{H5VL\_ATTR\_DELETE}, to delete an attribute on an object:
  \begin{enumerate}
  \item \texttt{char *attr\_name} (IN): the name of the attribute to delete.
  \end{enumerate}

\item \texttt{H5VL\_ATTR\_EXISTS}, to check if an attribute exists on a
  particular object specified in \texttt{obj}:
  \begin{enumerate}
  \item \texttt{char *attr\_name} (IN): the attribute name to check.
  \item \texttt{htri\_t *ret} (OUT): existence result, 0 if false, 1 if true.
  \end{enumerate}

\item \texttt{H5VL\_ATTR\_ITER}, to iterate over all attributes of an object and call a user-specified callback on each one:
  \begin{enumerate}
  \item \texttt{H5\_index\_t idx\_type} (IN): Type of index.\\
  \item \texttt{H5\_iter\_order\_t order} (IN): Order in which to iterate over index.\\
  \item \texttt{hsize\_t *idx} (IN/OUT): Initial and return offset within index.\\
  \item \texttt{H5A\_operator2\_t op} (IN): User-defined function to pass each attribute to. \\
  \item \texttt{void *op\_data} (IN/OUT): User data to pass through to and to be returned by iterator operator function. \\
  \end{enumerate}

\item \texttt{H5VL\_ATTR\_RENAME}, to rename an attribute on an object:
  \begin{enumerate}
  \item \texttt{char *old\_name} (IN): the original name of the attribute.
  \item \texttt{char *new\_name} (IN): the new name to assign for the attribute.
  \end{enumerate}

\end{itemize}

@@

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
    herr_t (*optional)(void *obj, H5VL_attr_optional_t opt_type, hid_t dxpl_id, void **req, va_list arguments);
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{13.5cm}}
  \texttt{obj} & (IN): The container or object where the operation needs to happen.\\
  \texttt{opt\_type} & (IN): The type of operation.\\
  \texttt{dxpl\_id} & (IN): The data transfer property list.\\
  \texttt{req} & (IN/OUT): A pointer to the asynchronous request of the operation created by the connector.\\
  \texttt{arguments} & (IN/OUT): argument list containing parameters and output pointers for the get operation. \\
\end{tabular}
\end{mdframed}

Each connector that requires connector-specific operations should be able to use the \texttt{opt\_type} parameter (a connector-specific enum/integer) to determine the type of the operation and parse the \texttt{va\_list arguments} appropriately.

@@

\begin{lstlisting}[caption={Structure for dataset callback routines, H5VLconnector.h}, captionpos=b, label={lst:DSclass}]
typedef struct H5VL_dataset_class_t {                                            
    void *(*create)(void *obj, const H5VL_loc_params_t *loc_params, const char *name, hid_t lcpl_id, hid_t type_id, hid_t space_id, hid_t dcpl_id, hid_t dapl_id, hid_t dxpl_id, void **req);
    void *(*open)(void *obj, const H5VL_loc_params_t *loc_params, const char *name, hid_t dapl_id, hid_t dxpl_id, void **req);
    herr_t (*read)(void *dset, hid_t mem_type_id, hid_t mem_space_id, hid_t file_space_id, hid_t dxpl_id, void * buf, void **req);
    herr_t (*write)(void *dset, hid_t mem_type_id, hid_t mem_space_id, hid_t file_space_id, hid_t dxpl_id, const void * buf, void **req);
    herr_t (*get)(void *obj, H5VL_dataset_get_t get_type, hid_t dxpl_id, void **req, va_list arguments);
    herr_t (*specific)(void *obj, H5VL_dataset_specific_t specific_type, hid_t dxpl_id, void **req, va_list arguments);
    herr_t (*optional)(void *obj, H5VL_dataset_optional_t opt_type, hid_t dxpl_id, void **req, va_list arguments);
    herr_t (*close) (void *dset, hid_t dxpl_id, void **req);
} H5VL_dataset_class_t; 
\end{lstlisting}

@@

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
    herr_t (*get)(void *dset, H5VL_dataset_get_t get_type, 
        hid_t dxpl_id, void **req, va_list arguments);
\end{lstlisting}

The \texttt{get\_type} argument is an \texttt{enum}:
\begin{lstlisting}
/* types for dataset GET callback */
typedef enum H5VL_dataset_get_t {                                                
    H5VL_DATASET_GET_DAPL,                  /* access property list                */
    H5VL_DATASET_GET_DCPL,                  /* creation property list              */
    H5VL_DATASET_GET_SPACE,                 /* dataspace                           */
    H5VL_DATASET_GET_SPACE_STATUS,          /* space  status                       */
    H5VL_DATASET_GET_STORAGE_SIZE,          /* storage size                        */
    H5VL_DATASET_GET_TYPE                   /* datatype                            */
} H5VL_dataset_get_t; 
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{13.5cm}}
  \texttt{dset} & (IN): The dataset object where information needs to be
  retrieved from.\\
  \texttt{get\_type} & (IN): The type of the information to retrieve.\\
  \texttt{dxpl\_id} & (IN): The data transfer property list.\\
  \texttt{req} & (IN/OUT): A pointer to the asynchronous request of the
  operation created by the connector.\\
  \texttt{arguments} & (IN/OUT): argument list containing parameters and
  output pointers for the get operation. \\
\end{tabular}
\end{mdframed}

The \texttt{arguments} argument contains a variable list of arguments
depending on the \texttt{get\_type} parameter. The following list shows
the argument list, in order, for each type:

\begin{itemize}
\item \texttt{H5VL\_DATASET\_GET\_DAPL}, to retrieve the dataset access
  property list of the dataset specified in \texttt{obj}:
  \begin{enumerate}
  \item \texttt{hid\_t *ret\_id} (OUT): buffer for the identifier of the
    dataset access property list.
  \end{enumerate}

\item \texttt{H5VL\_DATASET\_GET\_DCPL}, to retrieve the dataset creation
  property list of the dataset specified in \texttt{obj}:
  \begin{enumerate}
  \item \texttt{hid\_t *ret\_id} (OUT): buffer for the identifier of the
    dataset creation property list.
  \end{enumerate}

>>>>

\begin{lstlisting}[caption={VOL connector class, H5VLpublic.h},captionpos=b, label={lst:VOLclass}]
/* Class information for each VOL driver */
typedef struct H5VL_class_t {
    /* Overall connector fields & callbacks */
    unsigned           version;          /**< VOL connector class struct version #     */
    H5VL_class_value_t value;            /**< Value to identify connector              */
    const char *       name;             /**< Connector name (MUST be unique!)         */
    unsigned           conn_version;     /**< Version # of connector                   */
    unsigned           cap_flags;        /**< Capability flags for connector           */
    herr_t (*initialize)(hid_t vipl_id); /**< Connector initialization callback        */
    herr_t (*terminate)(void);           /**< Connector termination callback           */

    /* VOL framework */
    H5VL_info_class_t info_cls; /**< VOL info fields & callbacks  */
    H5VL_wrap_class_t wrap_cls; /**< VOL object wrap / retrieval callbacks */

    /* Data Model */
    H5VL_attr_class_t     attr_cls;     /**< Attribute (H5A*) class callbacks */
    H5VL_dataset_class_t  dataset_cls;  /**< Dataset (H5D*) class callbacks   */
    H5VL_datatype_class_t datatype_cls; /**< Datatype (H5T*) class callbacks  */
    H5VL_file_class_t     file_cls;     /**< File (H5F*) class callbacks      */
    H5VL_group_class_t    group_cls;    /**< Group (H5G*) class callbacks     */
    H5VL_link_class_t     link_cls;     /**< Link (H5L*) class callbacks      */
    H5VL_object_class_t   object_cls;   /**< Object (H5O*) class callbacks    */

    /* Infrastructure / Services */
    H5VL_introspect_class_t introspect_cls; /**< Container/connector introspection class callbacks */
    H5VL_request_class_t    request_cls;    /**< Asynchronous request class callbacks */
    H5VL_blob_class_t       blob_cls;       /**< 'Blob' class callbacks */
    H5VL_token_class_t      token_cls;      /**< VOL connector object token class callbacks */

    /* Catch-all */
    herr_t (*optional)(void *obj, H5VL_optional_args_t *args, hid_t dxpl_id,
                       void **req); /**< Optional callback */
} H5VL_class_t;

\end{lstlisting}

@@

The \texttt{value} field is a unique integer identifier that should be between 512 and 65535 for external, non-library connectors..

The \texttt{name} field is a string that uniquely identifies the VOL connector name.

The \texttt{conn\_version} is the connector version. This is currently not used by the library.

The \texttt{cap\_flags} holds bitwise capability/feature flags that determine which operations and capabilities are supported by a the VOL connector. These fields were enumerated in the previous section.

@@

The \texttt{terminate} field is a function pointer to a routine that a connector implements to terminate or finalize access to the connector. Implementing this function by the connector is not required since some connectors do not require any termination phase to the connector. In that case, the value of the function pointer should be set to NULL.

The rest of the fields in the \texttt{H5VL\_class\_t} struct are "subclasses'' that define all the VOL function callbacks that
are mapped to from the HDF5 API layer. Those subclasses are categorized into three categories, VOL Framework, Data Model, and Infrastructure / Services.

VOL Framework classes provide functionality for working with the VOL connectors themselves (e.g., working with connector strings) and with wrapping and unwrapping objects for passthrough connectors.

Data Model classes are those that provide functionality for accessing an HDF5 container and objects in that container as defined by the HDF5 data model.

Infrastructure / Service classes are those that provide services for users that are not related to the data model specifically. Asynchronous operations, for example, are a service that most connectors can implement, so we add a class for it in the VOL structure.

If a service becomes generic enough and common among many connectors, a class for it should be added to the VOL structure. However, many connectors can/will provide services that are not shared by other connectors. A good way to support these services is through an optional callback in the VOL structure which can be a hook from the API to the connector that provides those services, passing any necessary arguments needed without the HDF5 library having to worry about supporting that service. A similar API operation to allow users to use that service will be added. This API call would be similar to an “ioctl” call where any kind of operation can be supported and passed down to the connector that has enough knowledge from the user to interpret the type of the operation. All classes and their defined callbacks will be detailed in the following sub-sections.

To handle that large set of API routines, each class in the Data Model category has three generic callbacks, \texttt{get}, \texttt{specific}, and \texttt{optional} to handle the three set of API operations outline above respectively. To handle the varying parameters that can be passed to the callback, each callback will take a struct parameter that includes an enum {\tt get/specific} or integer {\tt optional} field indicating the operation and a union of the possible parameters {\tt get/specific} or void pointer to the parameters {\tt optional}.

The optional args struct used for all optional operations:
\begin{lstlisting}
/* Struct for all 'optional' callbacks */
typedef struct H5VL_optional_args_t {
    int   op_type; /* Operation to perform */
    void *args;    /* Pointer to operation's argument struct */
} H5VL_optional_args_t;
\end{lstlisting}

The {\tt opt\_type} member is the value assigned by the library when the optional operation was registered (or {\tt \#define}d in the case of the native VOL connector) and the {\tt args} member is a pointer to the optional operation's parameters (usually passed in as a struct).

Note that this differs from the HDF5 1.12.x scheme, which used {\tt va\_list}s.

The \texttt{optional} callback is a free for all callback where anything from the API layer is passed in directly. This callback is used to support connector specific operations in the API that other connectors should or would not know about. More information about types and the arguments for each type will be detailed in the corresponding class arguments.

@@

\begin{lstlisting}[caption={Structure for attribute callback routines, H5VLconnector.h}, captionpos=b, label={lst:ATTRclass}]
typedef struct H5VL_attr_class_t {                                               
    void *(*create)(void *obj, const H5VL_loc_params_t *loc_params, const char *attr_name, hid_t type_id, hid_t space_id, hid_t acpl_id, hid_t aapl_id, hid_t dxpl_id, void **req);
    void *(*open)(void *obj, const H5VL_loc_params_t *loc_params, const char *attr_name, hid_t aapl_id, hid_t dxpl_id, void **req);
    herr_t (*read)(void *attr, hid_t mem_type_id, void *buf, hid_t dxpl_id, void **req);
    herr_t (*write)(void *attr, hid_t mem_type_id, const void *buf, hid_t dxpl_id, void **req);
    herr_t (*get)(void *obj, H5VL_attr_get_args_t *args, hid_t dxpl_id, void **req);
    herr_t (*specific)(void *obj, const H5VL_loc_params_t *loc_params, H5VL_attr_specific_args_t *args, hid_t dxpl_id, void **req);            
    herr_t (*optional)(void *obj, H5VL_optional_args_t *args, hid_t dxpl_id, void **req); 
    herr_t (*close) (void *attr, hid_t dxpl_id, void **req);                     
} H5VL_attr_class_t; 
\end{lstlisting}

@@

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
    herr_t (*get)(void *obj, H5VL_attr_get_args_t *args, hid_t dxpl_id, 
        void **req);
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{13.5cm}}
  \texttt{obj} & (IN): An attribute or location object where information
  needs to be retrieved from.\\
  \texttt{args} & (IN/OUT): A pointer to the arguments struct.\\
  \texttt{dxpl\_id} & (IN): The data transfer property list.\\
  \texttt{req} & (IN/OUT): A pointer to the asynchronous request of the
  operation created by the connector.\\
\end{tabular}
\end{mdframed}

\begin{lstlisting}
/* Values for attribute 'get' operations */
typedef enum H5VL_attr_get_t {
    H5VL_ATTR_GET_ACPL,         /* creation property list              */
    H5VL_ATTR_GET_INFO,         /* info                                */
    H5VL_ATTR_GET_NAME,         /* access property list                */
    H5VL_ATTR_GET_SPACE,        /* dataspace                           */
    H5VL_ATTR_GET_STORAGE_SIZE, /* storage size                        */
    H5VL_ATTR_GET_TYPE          /* datatype                            */
} H5VL_attr_get_t;

/* Parameters for attribute 'get_name' operation */
typedef struct H5VL_attr_get_name_args_t {
    H5VL_loc_params_t loc_params;    /* Location parameters for object access */
    size_t            buf_size;      /* Size of attribute name buffer */
    char *            buf;           /* Buffer for attribute name (OUT) */
    size_t *          attr_name_len; /* Actual length of attribute name (OUT) */
} H5VL_attr_get_name_args_t;

/* Parameters for attribute 'get_info' operation */
typedef struct H5VL_attr_get_info_args_t {
    H5VL_loc_params_t loc_params; /* Location parameters for object access */
    const char *      attr_name;  /* Attribute name (for get_info_by_name) */
    H5A_info_t *      ainfo;      /* Attribute info (OUT) */
} H5VL_attr_get_info_args_t;

/* Parameters for attribute 'get' operations */
typedef struct H5VL_attr_get_args_t {
    H5VL_attr_get_t op_type; /* Operation to perform */

    /* Parameters for each operation */
    union {
        /* H5VL_ATTR_GET_ACPL */
        struct {
            hid_t acpl_id; /* Attribute creation property list ID (OUT) */
        } get_acpl;

        /* H5VL_ATTR_GET_INFO */
        H5VL_attr_get_info_args_t get_info; /* Attribute info */

        /* H5VL_ATTR_GET_NAME */
        H5VL_attr_get_name_args_t get_name; /* Attribute name */

        /* H5VL_ATTR_GET_SPACE */
        struct {
            hid_t space_id; /* Dataspace ID (OUT) */
        } get_space;

        /* H5VL_ATTR_GET_STORAGE_SIZE */
        struct {
            hsize_t *data_size; /* Size of attribute in file (OUT) */
        } get_storage_size;

        /* H5VL_ATTR_GET_TYPE */
        struct {
            hid_t type_id; /* Datatype ID (OUT) */
        } get_type;
    } args;
} H5VL_attr_get_args_t;
\end{lstlisting}

@@

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
	herr_t (*specific)(void *obj, H5VL_loc_params_t *loc_params, H5VL_attr_specific_args_t *args, hid_t dxpl_id, void **req);
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{13.5cm}}
  \texttt{obj} & (IN): The location object  where the operation needs to happen.\\
  \texttt{loc\_params} & (IN): A pointer to the location parameters as explained in Section~\ref{sec:map}.\\
  \texttt{args} & (IN/OUT): A pointer to the arguments struct.\\
  \texttt{dxpl\_id} & (IN): The data transfer property list.\\
  \texttt{req} & (IN/OUT): A pointer to the asynchronous request of the
  operation created by the connector.\\
\end{tabular}
\end{mdframed}

\begin{lstlisting}
/* Values for attribute 'specific' operation */
typedef enum H5VL_attr_specific_t {
    H5VL_ATTR_DELETE,        /* H5Adelete(_by_name)  */
    H5VL_ATTR_DELETE_BY_IDX, /* H5Adelete_by_idx     */
    H5VL_ATTR_EXISTS,        /* H5Aexists(_by_name)  */
    H5VL_ATTR_ITER,          /* H5Aiterate(_by_name) */
    H5VL_ATTR_RENAME         /* H5Arename(_by_name)  */
} H5VL_attr_specific_t;

/* Parameters for attribute 'iterate' operation */
typedef struct H5VL_attr_iterate_args_t {
    H5_index_t      idx_type; /* Type of index to iterate over */
    H5_iter_order_t order;    /* Order of index iteration */
    hsize_t *       idx;      /* Start/stop iteration index (IN/OUT) */
    H5A_operator2_t op;       /* Iteration callback function */
    void *          op_data;  /* Iteration callback context */
} H5VL_attr_iterate_args_t;

/* Parameters for attribute 'delete_by_idx' operation */
typedef struct H5VL_attr_delete_by_idx_args_t {
    H5_index_t      idx_type; /* Type of index to iterate over */
    H5_iter_order_t order;    /* Order of index iteration */
    hsize_t         n;        /* Iteration index */
} H5VL_attr_delete_by_idx_args_t;

/* Parameters for attribute 'specific' operations */
typedef struct H5VL_attr_specific_args_t {
    H5VL_attr_specific_t op_type; /* Operation to perform */

    /* Parameters for each operation */
    union {
        /* H5VL_ATTR_DELETE */
        struct {
            const char *name; /* Name of attribute to delete */
        } del;

        /* H5VL_ATTR_DELETE_BY_IDX */
        H5VL_attr_delete_by_idx_args_t delete_by_idx;

        /* H5VL_ATTR_EXISTS */
        struct {
            const char *name;   /* Name of attribute to check */
            hbool_t *   exists; /* Whether attribute exists (OUT) */
        } exists;

        /* H5VL_ATTR_ITER */
        H5VL_attr_iterate_args_t iterate;

        /* H5VL_ATTR_RENAME */
        struct {
            const char *old_name; /* Name of attribute to rename */
            const char *new_name; /* New attribute name */
        } rename;
    } args;
} H5VL_attr_specific_args_t;
\end{lstlisting}

@@

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
    herr_t (*optional)(void *obj, H5VL_optional_args_t *args, hid_t dxpl_id, void **req);
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{13.5cm}}
  \texttt{obj} & (IN): The container or object where the operation needs to happen.\\
  \texttt{args} & (IN/OUT): A pointer to the arguments struct.\\
  \texttt{dxpl\_id} & (IN): The data transfer property list.\\
  \texttt{req} & (IN/OUT): A pointer to the asynchronous request of the operation created by the connector.\\
\end{tabular}
\end{mdframed}

Each connector that requires connector-specific operations should compare the
value of the {\tt op\_type} field of the {\tt H5VL\_optional\_args\_t} struct
with the values returned from calling {\tt H5VLregister\_opt\_operation} to
determine how to handle the optional call and interpret the arguments passed
in the struct.

@@

\begin{lstlisting}[caption={Structure for dataset callback routines, H5VLconnector.h}, captionpos=b, label={lst:DSclass}]
typedef struct H5VL_dataset_class_t {                                            
    void *(*create)(void *obj, const H5VL_loc_params_t *loc_params, const char *name, hid_t lcpl_id, hid_t type_id, hid_t space_id, hid_t dcpl_id, hid_t dapl_id, hid_t dxpl_id, void **req);
    void *(*open)(void *obj, const H5VL_loc_params_t *loc_params, const char *name, hid_t dapl_id, hid_t dxpl_id, void **req);
    herr_t (*read)(void *dset, hid_t mem_type_id, hid_t mem_space_id, hid_t file_space_id, hid_t dxpl_id, void * buf, void **req);
    herr_t (*write)(void *dset, hid_t mem_type_id, hid_t mem_space_id, hid_t file_space_id, hid_t dxpl_id, const void * buf, void **req);
    herr_t (*get)(void *obj, H5VL_dataset_get_args_t *args, hid_t dxpl_id, void **req);
    herr_t (*specific)(void *obj, H5VL_dataset_specific_args_t *args, hid_t dxpl_id, void **req);
    herr_t (*optional)(void *obj, H5VL_optional_args_t *args, hid_t dxpl_id, void **req);
    herr_t (*close) (void *dset, hid_t dxpl_id, void **req);
} H5VL_dataset_class_t; 
\end{lstlisting}

@@

\begin{mdframed}[style=bgbox]
\textbf{Signature:}
\begin{lstlisting}
    herr_t (*get)(void *dset, H5VL_dataset_get_args_t *args, 
        hid_t dxpl_id, void **req);
\end{lstlisting}

\textbf{Arguments:}\\
\begin{tabular}{l p{13.5cm}}
  \texttt{dset} & (IN): The dataset object where information needs to be
  retrieved from.\\
  \texttt{args} & (IN/OUT): A pointer to the arguments struct.\\
  \texttt{dxpl\_id} & (IN): The data transfer property list.\\
  \texttt{req} & (IN/OUT): A pointer to the asynchronous request of the
  operation created by the connector.\\
\end{tabular}
\end{mdframed}

\begin{lstlisting}
/* Values for dataset 'get' operation */
typedef enum H5VL_dataset_get_t {
    H5VL_DATASET_GET_DAPL,         /* access property list                */
    H5VL_DATASET_GET_DCPL,         /* creation property list              */
    H5VL_DATASET_GET_SPACE,        /* dataspace                           */
    H5VL_DATASET_GET_SPACE_STATUS, /* space status                        */
    H5VL_DATASET_GET_STORAGE_SIZE, /* storage size                        */
    H5VL_DATASET_GET_TYPE          /* datatype                            */
} H5VL_dataset_get_t;

/* Parameters for dataset 'get' operations */
typedef struct H5VL_dataset_get_args_t {
    H5VL_dataset_get_t op_type; /* Operation to perform */

    /* Parameters for each operation */
    union {
        /* H5VL_DATASET_GET_DAPL */
        struct {
            hid_t dapl_id; /* Dataset access property list ID (OUT) */
        } get_dapl;

        /* H5VL_DATASET_GET_DCPL */
        struct {
            hid_t dcpl_id; /* Dataset creation property list ID (OUT) */
        } get_dcpl;

        /* H5VL_DATASET_GET_SPACE */
        struct {
            hid_t space_id; /* Dataspace ID (OUT) */
        } get_space;

        /* H5VL_DATASET_GET_SPACE_STATUS */
        struct {
            H5D_space_status_t *status; /* Storage space allocation status (OUT) */
        } get_space_status;

        /* H5VL_DATASET_GET_STORAGE_SIZE */
        struct {
            hsize_t *storage_size; /* Size of dataset's storage (OUT) */
        } get_storage_size;

        /* H5VL_DATASET_GET_TYPE */
        struct {
            hid_t type_id; /* Datatype ID (OUT) */
        } get_type;
    } args;
} H5VL_dataset_get_args_t;
\end{lstlisting}
